---
title: "Project 1 - Helicopters!"
author: "Jaskaran Singh, Michael Sarkis, Jack McCarthy"
date: "3/1/2022"
output: pdf_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(matrixStats)
library(kableExtra)
library(daewr)
library(readxl)
library(fdrtool)
knitr::opts_chunk$set(echo=TRUE)
```

# Introduction

The aim of this project is to optimize certain design parameters for the 
construction of a series of helicopter toys. The ideal result is a helicopter 
design that maximizes hang-time with minimal variance in the times across drops. 
However, as the budget is limited for the testing of various designs, we cannot 
perform more than 100 drops across all candidate designs. It is important that 
we perform these test drops in as economical a fashion as possible--that is, we 
would like to extract the most information possible about the effects of 
specific design decisions from each trial run as we can.

A factorial or fractional factorial experimental design is well-suited to a 
situation like this. Through this procedure, we may define factors with levels 
that represent different values of the design parameters of interest. Then, 
having collected data replicates for each specified combination of factor 
levels, we may perform an analysis to ascertain the effect of each factor along 
with the effects of possible interactions between factors. Knowing these effects 
will allow us to determine the optimal design having maximal hang-time and 
minimal variance.

# Design

A few test runs were performed to better understand how to make the helicopters 
and determined what changes to the design might impact the performance. Given 
the results from these tests, the following factors will considered in the 
construction of each candidate helicopter design:

|**Factor** |  **Description**  | **Low (-)** | **High (+)** |
|:---------:|-------------------|-------------|--------------|
|     A     | Total size        | 6.5"        | 11.0"        |
|     B     | Blade length      | 4.75"       | 5.75"        |
|     C     | Blade width       | 2.1"        | 3.0"         |
|     D     | Stabilizer length | 2.75"       | 4.75"        |
|     E     | Stabilizer width  | 0.5".        | 1.25"        |

For convenience we are referring to the bottom section of the helicopter that 
extends downward beneath the blades as the "stabilizer". We also noted during 
our initial tests that taping the bottom of the stabilizer is universally 
beneficial to the hang-time and stability of the helicopter. We will therefore 
being applying tape in all designs.

We have elected to conduct a $2_V^{5-1}$ fractional factorial experiment to 
analyze the effects of these factors on the hang-time. The generator for this 
experiment is $I=ABCDE$. This resulted in 16 distinct factor combinations with 5 
replicates for each combination, totaling 80 drops. The combinations and data 
are given in the following plot:

| **A** | **B** | **C** | **D** | **E** | 1 | 2 | 3 | 4 | 5 |
|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|:-:|
| - | - | - | - | + | 5.05 | 3.61 | 4.20 | 4.20 | 4.72 |
| - | - | - | + | - | 2.84 | 3.94 | 3.30 | 4.45 | 4.39 |
| - | - | + | - | - | 3.29 | 3.35 | 3.68 | 3.68 | 3.93 |
| - | - | + | + | + | 4.12 | 3.67 | 4.66 | 3.36 | 4.08 |
| - | + | - | - | - | 3.35 | 3.74 | 3.55 | 3.93 | 4.21 |
| - | + | - | + | + | 4.13 | 3.95 | 3.54 | 4.72 | 4.06 |
| - | + | + | - | + | 3.21 | 3.54 | 3.22 | 3.29 | 4.52 |
| - | + | + | + | - | 4.39 | 5.31 | 4.32 | 4.85 | 6.16 |
| + | - | - | - | - | 4.13 | 4.27 | 4.13 | 3.35 | 3.74 |
| + | - | - | + | + | 3.80 | 3.68 | 3.15 | 3.34 | 3.54 |
| + | - | + | - | + | 4.08 | 3.87 | 3.55 | 4.32 | 4.07 |
| + | - | + | + | - | 3.69 | 3.36 | 4.21 | 3.42 | 3.88 |
| + | + | - | - | + | 5.83 | 5.58 | 5.96 | 6.16 | 6.53 |
| + | + | - | + | - | 3.84 | 3.62 | 4.72 | 4.66 | 4.26 |
| + | + | + | - | - | 3.66 | 3.42 | 4.27 | 3.47 | 3.36 |
| + | + | + | + | + | 3.61 | 3.68 | 4.14 | 4.19 | 3.74 |

## Analysis

```{r, warning=F}
data <- read_excel("flight_data.xlsx") %>%
  pivot_longer(r1:r5, names_to="rep", values_to="time") %>%
  mutate_at(.funs=funs(ifelse(.=='+', 1, -1)),
            .vars=vars(A:E)) %>%
  select(-rep)
```

```{r}
# calculate main location effect
main.loc <- function(df, fac, resp) {
  # for column selection
  fac  <- sym(fac)
  resp <- enquo(resp)
  
  # calculate main effect
  df %>%
    group_by_at(1:5) %>%
    summarise(y.bar=mean(!!resp)) %>%
    group_by(!!fac) %>%
    summarise(mean.y.bar=mean(y.bar)) %>%
    mutate(to.add=!!fac*mean.y.bar) %>%
    pull(to.add) %>% sum()
}

main.disp <- function(df, fac, resp) {
  # for column selection
  fac  <- sym(fac)
  resp <- enquo(resp)
  
  # calculate main effect
  df %>%
    group_by_at(1:5) %>%
    summarise(lns.2=log(var(!!resp))) %>%
    group_by(!!fac) %>%
    summarise(mean.lns.2=mean(lns.2)) %>%
    mutate(to.add=!!fac*mean.lns.2) %>%
    pull(to.add) %>% sum()
}

two.fac.loc <- function(df, fac, resp) {
  facs <- strsplit(fac, "")[[1]]
  fac1 <- sym(facs[1])
  fac2 <- sym(facs[2])
  resp <- enquo(resp)
  
  # calculate two-factor location
  eff.1 <- df %>%
    filter(!!fac2>0) %>%
    main.loc(., facs[1], !!resp)
  eff.2 <- df %>%
    filter(!!fac2<0) %>%
    main.loc(., facs[1], !!resp)
  0.5 * (eff.1 - eff.2)
}

two.fac.disp <- function(df, fac, resp) {
  facs <- strsplit(fac, "")[[1]]
  fac1 <- sym(facs[1])
  fac2 <- sym(facs[2])
  resp <- enquo(resp)
  
  # calculate two-factor dispersion
  eff.1 <- df %>%
    filter(!!fac2>0) %>%
    main.disp(., facs[1], !!resp)
  eff.2 <- df %>%
    filter(!!fac2<0) %>%
    main.disp(., facs[1], !!resp)
  0.5 * (eff.1 - eff.2)
}
```

```{r, message=F, warning=F}
# function wrappers
flight.main.loc     <- function(fac) { main.loc(data, fac, time) }
flight.main.disp    <- function(fac) { main.disp(data, fac, time) }
flight.two.fac.loc  <- function(fac) { two.fac.loc(data, fac, time) }
flight.two.fac.disp <- function(fac) { two.fac.disp(data, fac, time) }

# effect labels
labels=c("A",  "B",  "C",  "D",  "E", 
         "AB", "AC", "AD", "AE", "BC", 
         "BD", "BE", "CD", "CE", "DE")

# effect table
eff.tbl <- tibble(
  label    = labels,
  loc.eff  = c(sapply(labels[1:5],   flight.main.loc),
               sapply(labels[6:10],  flight.two.fac.loc),
               sapply(labels[11:15], flight.two.fac.loc)),
  disp.eff = c(sapply(labels[1:5],   flight.main.disp),
               sapply(labels[6:10],  flight.two.fac.disp),
               sapply(labels[11:15], flight.two.fac.disp))
)
```

```{r, echo=F}
eff.tbl %>%
  mutate(sgn=as.factor(sign(loc.eff)),
         loc.eff=abs(loc.eff)) %>%
  arrange(loc.eff) %>%
  mutate(i=1:nrow(.)) %>%
  mutate(z=qnorm(((i - 0.5)/nrow(.) + 1)/2)) %>%
  ggplot(aes(x=z, y=loc.eff)) +
    geom_point(aes(color=sgn)) +
    geom_text(aes(label=label),
              hjust=1, vjust=-0.5) +
    labs(x="Half-Normal Z-Scores",
         y="Absolute Location Effects",
         title="Location Effects Half-Normal Plot",
         color="Sign of Effect") +
    scale_color_discrete(labels=c("-", "+")) +
    theme_bw()

eff.tbl %>%
  mutate(sgn=as.factor(sign(disp.eff)),
         disp.eff=abs(disp.eff)) %>%
  arrange(disp.eff) %>%
  mutate(i=1:nrow(.)) %>%
  mutate(z=qnorm(((i - 0.5)/nrow(.) + 1)/2)) %>%
  ggplot(aes(x=z, y=disp.eff)) +
    geom_point(aes(color=sgn)) +
    geom_text(aes(label=label),
              hjust=1, vjust=-0.5) +
    labs(x="Half-Normal Z-Scores",
         y="Absolute Dispersion Effects",
         title="Dispersion Effects Half-Normal Plot",
         color="Sign of Effect") +
    scale_color_discrete(labels=c("-", "+")) +
    theme_bw()
```

```{r}
loc.all <- abs(eff.tbl$loc.eff)
cutoff <- 3.75 * median(loc.all)
int.sub <- loc.all[loc.all <= cutoff]
pse <- 1.5 * median(int.sub)

lenth.med <- eff.tbl %>%
  mutate(loc.abs=abs(loc.eff),
         disp.abs=abs(disp.eff)) %>%
  mutate(loc.cut=3.75*median(loc.abs),
         disp.cut=3.75*median(disp.abs)) %>%
  mutate(loc.inc=ifelse(loc.abs<=loc.cut, T, F),
         disp.inc=ifelse(disp.abs<=disp.cut, T, F)) %>%
  filter(loc.inc, disp.inc) %>%
  summarise(loc.pse=1.5*median(loc.abs),
            disp.pse=1.5*median(disp.abs)) %>%
  mutate(ier=2.16)

lenth.tbl <- bind_cols(eff.tbl, lenth.med) %>%
  mutate(loc.sig=abs(loc.eff)/loc.pse > ier,
         disp.sig=abs(disp.eff)/disp.pse > ier) %>%
  mutate(label=labels) %>%
  select(label, loc.sig, disp.sig)

lenth.tbl %>% kable(col.names=c("Factors", 
                                "Loc. Significant?", 
                                "Disp. Significant?"),
                    align=c('l', 'c', 'c')) %>%
  kable_styling(position="center") %>%
  row_spec(0, bold=T)
```

Testing for significance via Lenth's method, we find that there are no 
significant effects for location, but **A** and **DE** are significant for 
dispersion as we originally surmized from the half-normal plots.

## Optimization

An optimal design is one that stays in the air as long as possible with 
consistency between drops. In more rigorous terms, we want a design that 
maximizes the expected hang-time while minimizing dispersion.

